RNApyro

### 2012-10-05 Recomb/methods_RECOMB.tex
Initial commit

%!TEX root = main_RECOMB.tex
\section{Methods}
\label{sec:methods}
Let $\Omega$ be an un-gapped RNA alignment, $S$ its associated secondary structure, $s$ an RNA 
sequence and $m\geq 0$. $S$ is considered as  one derivation of the SCFG generating  
all  secondary structures of length  $|s|$. We are interested in  
the probability of a given position being a specific nucleotides,
over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$ 
(formally 
$\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
We define as a variant of the
 \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
 desired probability,  the two  functions
$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former presented in 
Eq.~\ref{eq:Z_in} and Eq.~\ref{eq:Z_rec}, our 
version of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing 
$m$ mutations. The latter in Eq.~\ref{eq:Y_in} and Eq.\ref{eq:Y_rec},
 computes the \emph{outside}, in particular,  
the partition function considering only between $[0,i]\cup[j,n-1]$
knowing  that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing 
 $m$ mutations outside.

The Boltzmann weights are a combination of the base pairs stacking energy, 
using as values those of the NNDB~\cite{Turner2010}, and
the average isostericity difference
between the mutant and $\Omega$, and $s$ with $\Omega$, using the isostericity values
 as defined in~\cite{Stombaugh2009}. The value of $10$ was used for the isostericity
 of any base pair 
 compared with $GG$ given that the latter base pair was not
  found in the~\cite{Stombaugh2009} tables,
 and the range of values is $[0,9.7]$, $0$ for a perfect isostericity.

>>>>

%!TEX root = main_RECOMB.tex
\section{Methods}
\label{sec:methods}

\subsection{Model}
Our probabilistic model aims at capturing both the stability of the folded RNA and its ability to adopt a predefined 3D conformation.
To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function which combines free-energy and isostericity contributions.

The free-energy contribution corresponds to the stacking base-pairs free-erngies of the Turner XXXXX model, as retrieved from the NNDB~\cite{Turner2010}.
This simplification of the Turner energy .
The imprecision introduced by our simplification of the Turner model not only
greatly eases the design of dynamic-programming equations but  also causes 
a limited imprecision since the targeted secondary structure is fixed 
(e.g. multiloops do not account for base-specific contributions).

The isostericity term captures the superimposability of corresponding base-pairs in the initial and mutated structures,
%$\Omega$, and $s$ with $\Omega$, 
using isostericity values defined in~\cite{Stombaugh2009}. 
The range of values is $[0,9.7]$, with $0$ standing for a perfect isostericity, 
and $10$ being used for missing values in Stombaugh~\emph{et al}~\cite{Stombaugh2009} tables.
The relative contribution of these terms in the pseudo-energy function is then adjusted using a parameter $\alpha\in[0,1]$.

Let $\Omega$ be an un-gapped RNA alignment, $S$ its associated secondary structure, $s$ an RNA 
sequence and $m\geq 0$. $S$ is considered as  one derivation of the SCFG generating  
all  secondary structures of length  $|s|$. We are interested in  
the probability of a given position being a specific nucleotides,
over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$ 
(formally 
$\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
We define as a variant of the
 \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
 desired probability,  the two  functions
$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former presented in 
Eq.~\ref{eq:Z_in} and Eq.~\ref{eq:Z_rec}, our 
version of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing 
$m$ mutations. The latter in Eq.~\ref{eq:Y_in} and Eq.\ref{eq:Y_rec},
 computes the \emph{outside}, in particular,  
the partition function considering only between $[0,i]\cup[j,n-1]$
knowing  that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing 
 $m$ mutations outside.
